---
title: "TestMarkdown-Analysis"
output:
  pdf_document: default
  html_document: default
date: "`r Sys.Date()`"
# YAML-MetaData
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Software Requirements:


```{r eval=FALSE, include=TRUE}
- R Installation
- RStudio for IDE
- LaTeX Installation for rendering RMarkdown to PDF (this file) 

- GitHub Account for Repository/Data access

```


## Setting up the Environment:


### Required R Packages:

```{r eval=FALSE, include=TRUE}
install.packages("rmarkdown")
install.packages("devtools")

install.packages("tidyverse")
install.packages("metafor")
install.packages("meta")

install.packages("tidyr")

install.packages("weightr")

```



### Loading first Packages:

```{r echo=TRUE}
library(rmarkdown)

library(tidyverse)


library(metafor)
library(meta)
library(dmetar)

library (dplyr)
library(tidyr)

library(stringr)
library(readr)

```


### Installing further R Packages:

```{r eval=FALSE, include=TRUE}
devtools::install_github("MathiasHarrer/dmetar") 


```



### Loading remaining Packages:

```{r echo=TRUE}
library(dmetar)

```




### Cloning Repository from Git-Hub (Fetching Data):
```{r eval=FALSE, include=TRUE}
usethis::create_from_github(
  "https://github.com/VJMeyer/HPVPre_Repo.git",
  destdir = "~/path/to/where/you/want/the/local/repo/"
)

```


### Set Working-Directory:

```{r eval=FALSE, include=TRUE}

setwd("~/path/to/where/you/want/the/local/repo/")

```


### Set Private Working-Directory:

```{r eval=TRUE, include=FALSE}
setwd("/Users/vjm/Library/Mobile Documents/iCloud~md~obsidian/Documents/Vaulty/1_Foundations/Together/MubsThesisLocal/RStudioGitMirror")

```



### Confirm Working-Directory:

```{r eval=TRUE, include=FALSE}
#confirm working directory
getwd()
list.files() #list files within it

```




### Load Data:

```{r Loading Data, echo=TRUE}
library(tidyverse)

any_type <- read_csv("Anytype_OLDI.csv")  
hr_type <- read_csv("HR_OLDI.csv") 
gtype <- read_csv("gtype.csv") 

```

# Quantitative Analysis

## Data Cleaning

### Ensure there are no MISSING values in the important columns of ANYTYPE Data:

```{r , echo=TRUE}
library(tidyverse)


class(any_type)
head(any_type)
str(any_type)
summary(any_type)
view(any_type)

```



### Filter rows where num_older_wom is NA and select study_id

```{r echo=TRUE}
na_in_num_older_wom <- any_type %>%
  filter(is.na(num_older_wom)) %>%
  select(study_id)
```

### Filter rows where num_hpv_pos is NA and select study_id

```{r }
na_in_num_hpv_pos <-any_type %>%
  filter(is.na(num_hpv_pos)) %>%
  select(study_id)
```

### Filter rows where num_hpv_neg is NA and select study_id

```{r }
na_in_num_hpv_neg <- any_type %>%
  filter(is.na(num_hpv_neg)) %>%
  select(study_id)
```

### View the results

```{r }
na_in_num_older_wom
na_in_num_hpv_pos
na_in_num_hpv_neg
```

there was missing data in one row (63O) of num_hpv_neg which was corrected 

Filter out rows where the identifier ends with 'P'

```{r }
any_type_filtered <- any_type %>%
  filter(!str_detect(study_id, "P$"))

summary(any_type_filtered)

nrow(any_type_filtered)
```



### Extract and view unique study ids in hr_type_filtered to make sure it is correct

```{r }
unique_study_ids_any <- any_type_filtered %>%
  select(study_id) %>%
  distinct()

print(unique_study_ids_any, n = 83)
```



###Ensure there are no MISSING values in the columns  of  HR TYPE  where there should be data

```{r}
class(hr_type)
head(hr_type)
str(hr_type)
summary(hr_type)
view(hr_type)
nrow(hr_type)
```

### Filter rows where num_older_wom is NA and select study_id

```{r}
na_in_num_older_wom_hr <- hr_type %>%
  filter(is.na(num_older_wom)) %>%
  select(study_id)
```

### Filter rows where num_hr_hpv_pos is NA and select study_id

```{r}
na_in_num_hr_hpv_pos <- hr_type %>%
  filter(is.na(num_hr_hpv_pos)) %>%
  select(study_id)
```

### Filter rows where hr_prev is NA and select study_id

```{r}
na_in_hr_prev <- hr_type %>%
  filter(is.na(hr_prev)) %>%
  select(study_id)
```

### View the results

```{r}
na_in_num_older_wom_hr
na_in_num_hr_hpv_pos
na_in_hr_prev
```

we must exclude study id 23HG from the HR analysis because it has crude numbers for all gtypes but not for HR type
Filter out rows where the identifier ends with 'P' or is '23HG'

```{r}
hr_type_filtered <- hr_type %>%
  filter(!str_detect(study_id, "P$"), study_id != "23HG")


summary(hr_type_filtered)
nrow(hr_type_filtered)



```

### Extract and view unique study ids in hr_type_filtered to make sure it is correct

```{r}
unique_study_ids_hr <- hr_type_filtered %>%
  select(study_id) %>%
  distinct()

print(unique_study_ids_hr, n = 112)
```



### Filter out rows where the identifier ends with 'P' for prevalance only for gtype_filtered 

###trim whitespaces

```{r}
gtype <- gtype %>%
  mutate(study_id = trimws(study_id))

gtype_filtered <- gtype %>%
  filter(!str_detect(study_id, "P$"))

class(gtype_filtered)
head(gtype_filtered)
str(gtype_filtered)
summary(gtype_filtered)
view(gtype_filtered)

summary(gtype_filtered)
nrow(gtype_filtered)
```



### Extract and view unique study ids in gtype_filtered to make sure it is correct

```{r}
unique_study_ids_gtype <- gtype_filtered %>%
  select(study_id) %>%
  distinct()

nrow(unique_study_ids_gtype)
print(unique_study_ids_gtype, n = 43)

```

```{r}
#Data frames for quantitative analysis ----
# any hpv type analysis ---> any_type_filtered
#hr type analysis ---> hr_type_filtered
#type-specific analysis --->  gtype_filtered



```



 


### Meta-analysis model for any_type_filtered ----


```{r}
summary(any_type_filtered) # there are zero event values in the num_hpv_pos
```

### Identifying zero-event cases

```{r}
zero_event_cases <- any_type_filtered[any_type_filtered$num_hpv_pos == 0, ]
```

### Counting the number of zero-event cases

```{r}
num_zero_event_cases <- nrow(zero_event_cases)
```

### Printing the number of zero-event cases

```{r}
print(paste("Number of zero-event cases:", num_zero_event_cases))
```

### Printing the study IDs of zero-event cases

```{r}
if(num_zero_event_cases > 0) {
  print("Study IDs with zero-event cases:")
  print(zero_event_cases$author_year)
  print(zero_event_cases$study_id)
} else {
  print("There are no zero-event cases.")
}
```



### Random effects MA, Freeman-Tukey double arcsine transformation, restricted maximum likelihood estimator, Knapp-Hartung adjustment
```{r}
any_type_ma <- metaprop(event = num_hpv_pos, 
                        n = num_older_wom,
                        studlab = author_year,
                        data = any_type_filtered, 
                        sm = "PFT", 
                        method.tau = "REML",
                        method.ci = "NAsm",### Specify the method for confidence intervals
                        add = 0,
                        fixed = FALSE,
                        random = TRUE,
                        hakn = TRUE,
                        title = "Anytype HPV Prevalence in Women (50+) with Predominantly Normal Cytology")

summary(any_type_ma)
```



# ANYTYPE FOREST PLOT ----

## Create a forest plot based on the meta-analysis results
```{r}
pdf("forestplot.pdf", width=10, height=20)  # Width and height in inches, adjust as needed

forest(any_type_ma, 
       common = TRUE, 
       print.tau2 = TRUE, 
       print.Q = TRUE, 
       print.pval.Q = TRUE, 
       print.I2 = TRUE, 
       rightcols = FALSE, 
       pooled.totals = TRUE, 
       weight.study = "random", 
       leftcols = c("studlab", "num_hpv_pos", "num_older_wom", "effect", "ci"), 
       leftlabs = c("Study", "HPV+ Cases", "Total", "Prevalence", "95% C.I."), 
       xlab = "Pooled Prevalence Rate", 
       smlab = "", 
       xlim = c(0,1), 
       pscale = 1, 
       squaresize = 0.5, 
       fs.hetstat = 10, 
       digits = 2, 
       col.square = "navy", 
       col.square.lines = "navy", 
       col.diamond = "maroon", 
       col.diamond.lines = "maroon")

dev.off()
```

```{r}
pdf("forestplot2.pdf", width=10, height=20)  # Width and height in inches, adjust as needed

forest(any_type_ma,
       common = TRUE, 
       print.tau2 = TRUE, 
       print.Q = TRUE, 
       print.pval.Q = TRUE, 
       print.I2 = TRUE, 
       rightcols = FALSE, 
       pooled.totals = TRUE,
       weight.study = "random", 
       xlab = "Pooled Prevalence Rate",
       leftcols = c("studlab", "num_hpv_pos", "n", "effect", "ci"), 
       leftlabs = c("Study", "HPV+ Cases", "Total", "Prevalence", "95% C.I."), 
       smlab = "", 
       xlim = c(0,1), 
       pscale = 1, 
       squaresize = 0.5, 
       fs.hetstat = 10, 
       digits = 2, 
       col.square = "navy", 
       col.square.lines = "navy", 
       col.diamond = "maroon", 
       col.diamond.lines = "maroon"
)


dev.off()
```

## Show Forest Plot
```{r}
system2('open', args = 'forestplot2.pdf', wait = FALSE)
```


# Funnel Plot
```{r}
funnel(any_type_ma, main="Funnel Plot for Meta-Analysis", xlim=c(0, 1))

```




















